---
title: "Importing and Scraping Data ...   \n and random tidying tools"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

## Setup

Welcome to another tutorial for this class, COMP/STAT 112: *Introduction to Data Science*! It will be similar to the others, including demo videos and files embedded in this document and practice problems with hints or solutions at the end. There are some new libraries, so be sure to install those first.

As most of our files do, we start this one with three R code chunks: 1. options, 2. libraries and settings, 3. data. 

```{r setup}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

```{r libraries}
library(tidyverse)     # for data cleaning and plotting
library(googlesheets4) # for reading googlesheet data
library(lubridate)     # for date manipulation
library(openintro)     # for the abbr2state() function
library(palmerpenguins)# for Palmer penguin data
library(maps)          # for map data
library(ggmap)         # for mapping points on maps
library(gplots)        # for col2hex() function
library(RColorBrewer)  # for color palettes
library(sf)            # for working with spatial data
library(leaflet)       # for highly customizable mapping
library(ggthemes)      # for more themes (including theme_map())
library(plotly)        # for the ggplotly() - basic interactivity
library(gganimate)     # for adding animation layers to ggplots
library(gifski)        # for creating the gif (don't need to load this library every time,but need it installed)
library(transformr)    # for "tweening" (gganimate)
library(shiny)         # for creating interactive apps
library(patchwork)     # for nicely combining ggplot2 graphs  
library(gt)            # for creating nice tables
library(knitr)         # for the kable() function for creating nice tables
library(rvest)         # for scraping data
library(robotstxt)     # for checking if you can scrape data
gs4_deauth()           # To not have to authorize each time you knit.
theme_set(theme_minimal())
```

```{r my_libraries, include=FALSE}
# Lisa needs this, students don't
library(downloadthis) # for including download buttons for files
library(flair)        # for highlighting code
library(xaringanExtra)# for small slides and other cool things
```

```{r data}
# Lisa's garden data
garden_harvest <- read_sheet("https://docs.google.com/spreadsheets/d/1DekSazCzKqPS2jnGhKue7tLxRU3GVL1oxi-4bEM5IWw/edit?usp=sharing") %>% 
  mutate(date = ymd(date))
```

## Learning Goals

After this tutorial, you should be able to do the following:

* Import data into R that is stored in a common file type (.csv, .txt, excel, etc) or in a Google spreadsheet.  

* Find resources to read in data that is in a format other than one of the more common formats.

* Use `rvest()` functions to scrape data from a simple webpage and recognize when scraping the data will require more advanced tools.

* Create nice tables with the `knitr` `kable()` functions and the `gt` functions. 

* Use `patchwork` to display related plots together nicely.

## Importing Data

### Common Import functions

### Using the Import Dataset Wizard

### Resources

* [R4DS](https://r4ds.had.co.nz/data-import.html)

* [readr documentation](https://readr.tidyverse.org/) and data import cheatsheet

## Scraping Data


